SunPy - Python for Solar Physics

### 2014-01-28 3-4-Helioviewer.tex
small spacing edit to Helioviewer file
Some small typo fixed and latex stuff

SunPy provides the ability to download images hosted by the
Helioviewer Project.  The aim of the Helioviewer Project is to enable
the exploration of solar and heliospheric data from multiple data
sources (such as instrumentation and feature/event catalogs) via
easy-to-use visual interfaces.  The primary data provided by the
Helioviewer Project are images in the JPEG2000 format (for further
details on the Helioviewer Project see \cite{muller2009}). The
JPEG2000 files are typically highly compressed compared to the source
FITS files they are generated from, and so can be a used as a shortcut
to visualize large amounts of data from multiple sources.  SunPy is
also used in Helioviewer production servers to manage the download and
ingestion of JPEG2000 files from remote servers.

The Helioviewer Project categorizes image data based on the physical
construction of the source instrument, using a simple hierarchy:
observatory $\rightarrow$ instrument $\rightarrow$ detector
$\rightarrow$ measurement, where '$\rightarrow$' means 'provides
(possibly multiple)'.  Each Helioviewer Project JPEG2000 file contains
metadata which are based (in part) on the original FITS header
information, and carry sufficient information to permit overlay with
other Helioviewer JPEG2000 files. Images can be accessed either as
PNGs (Section \ref{sssec:hv:png} or as JPEG2000 files (Section
\ref{sssec:hv:jp}).

@@

Listing \ref{code:hv:downloadlatestpng} demonstrates how to download a
PNG image of the latest AIA 304\AA\ image available on
\url{Helioviewer.org}.

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy.net.helioviewer import HelioviewerClient
>>> hv = HelioviewerClient()
>>> hv.download_png('2099/01/01', 4.8, "[SDO,AIA,AIA,304,1,100]",
...                 x0=0, y0=0, width=512, height=512)
\end{minted}
\includegraphics[width=0.8\columnwidth]{helioviewer_latest_aia_304}
\caption{Acquisition of a PNG file showing the latest AIA 304\AA\ image 
available at \url{www.helioviewer.org}.}
\label{code:hv:downloadlatestpng}
\end{listing}

The first argument is the requested time of the image.  Helioviewer
selects images closest to the requested time.  In this case, the
requested time is in the future and so Helioviewer will find the most
recent available image.  The second argument refers to the image
resolution in arcseconds per pixel (larger values mean lower
resolution).  The third argument is a comma-delimited string detailing
the requested observatory $\rightarrow$ instrument $\rightarrow$
detector $\rightarrow$ measurement combination.  The final two numbers
in this string are the visibility and the opacity of the this image
layer (1/0 is visible/invisible with opacity in the range
$0\rightarrow100$, with $100$ meaning fully opaque).  The quantities
$x0$ and $y0$ are the $x$ and $y$ center points about which to center
the image (measured in helio-projective cartesian co-ordinates) and
the \texttt{width} and \texttt{height} are the pixel values for the
image dimensions.

@@

The principle difference between this code snippet and Listing
\ref{code:hv:downloadlatestpng} is in the layer string (the third
argument).  The layer string has three separate comma separated
entries detailing the data source, its visibility and its opacity.
This functionality makes it simple for SunPy users to generate complex
images from multiple, correctly overlaid, image data sources.

@@

As noted above, Helioviewer JPEG2000 files contain metadata that allow
positioning of the image data.  There is sufficient metadata in each
file to permit the creation of a SunPy map object (see Section
\ref{ssec:maps}) from a Helioviewer JPEG2000 file.  This allows image
data to be manipulated in the same way as any other map object.

@@

Listing \ref{code:downloadjp2} demonstrates the querying, downloading,
reading and conversion of a Helioviewer JPEG2000 file into a SunPy Map
object.  This functionality allows users to visualize and manipulate
Helioviewer-supplied image data in an identical fashion to a SunPy map
object generated from FITS data (see Section \ref{sec:maps}).

>>>>

SunPy provides the ability to download images hosted by the
Helioviewer Project (\url{http://helioviewer.org}).  
The aim of the Helioviewer Project is to enable
the exploration of solar and heliospheric data from multiple data
sources (such as instrumentation and feature/event catalogues) via
easy-to-use visual interfaces.  The primary data provided by the
Helioviewer Project are images in the JPEG2000 format (for further
details on the Helioviewer Project see \cite{muller2009}). The
JPEG2000 files are typically highly compressed compared to the source
FITS files they are generated from, and so can be a used as a shortcut
to visualise large amounts of data from multiple sources.  SunPy is
also used in Helioviewer production servers to manage the download and
ingestion of JPEG2000 files from remote servers.

The Helioviewer Project categorises image data based on the physical
construction of the source instrument, using a simple hierarchy:
observatory $\rightarrow$ instrument $\rightarrow$ detector
$\rightarrow$ measurement, where '$\rightarrow$' means 'provides
(possibly multiple)'.  Each Helioviewer Project JPEG2000 file contains
metadata which are based (in part) on the original FITS header
information, and carry sufficient information to permit overlay with
other Helioviewer JPEG2000 files. Images can be accessed either as
PNGs (Section \ref{sssec:hv:png}) or as JPEG2000 files (Section
\ref{sssec:hv:jp}).

@@

Listing \ref{code:hv:downloadlatestpng} demonstrates how to download a
PNG image of the latest AIA 304$\,$\AA\ image available on
\url{Helioviewer.org}.

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy.net.helioviewer import HelioviewerClient
>>> hv = HelioviewerClient()
>>> hv.download_png('2099/01/01', 4.8, "[SDO,AIA,AIA,304,1,100]",
...                 x0=0, y0=0, width=512, height=512)
\end{minted}
\includegraphics[width=0.8\columnwidth]{helioviewer_latest_aia_304}
%centered image?
\caption{Acquisition of a PNG file showing the latest AIA 304\AA\ image 
available at \url{www.helioviewer.org}.}
\label{code:hv:downloadlatestpng}
\end{listing}

The first argument is the requested time of the image.  Helioviewer
selects images closest to the requested time.  In this case, the
requested time is in the future and so Helioviewer will find the most
recent available image.  The second argument refers to the image
resolution in arcseconds per pixel (larger values mean lower
resolution).  The third argument is a comma-delimited string detailing
the requested observatory $\rightarrow$ instrument $\rightarrow$
detector $\rightarrow$ measurement combination.  The final two numbers
in this string are the visibility and the opacity of the this image
layer (1/0 is visible/invisible with opacity in the range
$0\rightarrow100$, with $100$ meaning fully opaque).  The quantities
$x0$ and $y0$ are the $x$ and $y$ centre points about which to centre
the image (measured in helio-projective cartesian co-ordinates) and
the \texttt{width} and \texttt{height} are the pixel values for the
image dimensions.

@@

The principle difference between Listings~\ref{code:hv:overlaid} and 
\ref{code:hv:downloadlatestpng} is in the layer string (the third
argument).  The layer string has three separate comma separated
entries detailing the data source, its visibility and its opacity.
This functionality makes it simple for SunPy users to generate complex
images from multiple, correctly overlaid, image data sources.

@@

As noted above, Helioviewer JPEG2000 files contain metadata that allow
positioning of the image data.  There is sufficient metadata in each
file to permit the creation of a SunPy map object (see Section
\ref{ssec:map}) from a Helioviewer JPEG2000 file.  This allows image
data to be manipulated in the same way as any other map object.

@@

Listing \ref{code:downloadjp2} demonstrates the querying, downloading,
reading and conversion of a Helioviewer JPEG2000 file into a SunPy Map
object.  This functionality allows users to visualise and manipulate
Helioviewer-supplied image data in an identical fashion to a SunPy map
object generated from FITS data (see Section \ref{sec:map}).

### 2014-01-28 3-5-Database.tex
Some small typo fixed and latex stuff

\subsection{The File Database}
Easy access to large quantities of solar data frequently leads to datafiles accumulating
in local storage such as laptops and desktop computers. Keeping data organized and available
is typically a cumbersome task for the average user. The file database is a sub-package of 
SunPy which is meant to solve this problem by providing a unified database to store and manage information about 
local data files. The database package was developed as part of the 2013 
Google Summer of Code.

The database sub-package makes use of a database supported by
SQLAlchemy(\url{http://www.sqlalchemy.org}). This library was chosen
since it supports many SQL dialects. If SQLite is choosen, the database is stored as a single file which is
created automatically. This solution makes it possible to easily backup, copy and distribute
a database file. A server-based database, on the other hand, could be used by a collaboration 
who work together from different machines on the
same data: a central database server stores all data and the clients connect to
it to read or write data.

The database can store and manage all data that is retrieved via sunpy from the net
package though currently only the \textsc{VSO} package is supported. 
It also possible to add entries manually by file or directory. The package also provides
a \texttt{fetch} method which provides a unified data search which includes both local files
and files on the \textsc{VSO} are searched. This reduces the likelihood of downloading the same file 
multiple times. When a file is added to the database the file is scanned and a file hash is produced. 
The current date is associated with the entry along with metadata summaries such 
as instrument, date of observation, field of view, etc. 
The database also provides the ability to associate custom meta-data to 
each database entry such as keywords, comments, and favoriting.

@@

\begin{listing}[H]
\begin{minted}{python}
>>> from sunpy.net import vso
>>> from sunpy.database import Database
>>> database = Database('sqlite:///')
>>> database.download(
...     vso.attrs.Time((2011, 9, 20, 1), (2011, 9, 20, 2)),
...     vso.attrs.Instrument('RHESSI'))
>>> len(database)
6
>>> from sunpy.database.tables import display_entries
>>> print display_entries(
...     database,
...     ['id', 'observation_time_start', 'wavemin', 'wavemax', 'download_time'])
id observation_time_start wavemin        wavemax           download_time      
-- ---------------------- -------        -------           -------------      
1  2011-09-20 01:09:20    0.413280643067 7.29318781883e-05 2014-01-18 19:21:35

>>>>

\subsection{The File Database}\label{ssec:db}

Easy access to large quantities of solar data frequently leads to datafiles accumulating
in local storage such as laptops and desktop computers. Keeping data organised and available
is typically a cumbersome task for the average user. The file database is a sub-package of 
SunPy which is meant to solve this problem by providing a unified database to store and manage information about 
local data files. The database package was developed as part of GSOC 2013.

The database sub-package makes use of a database supported by
SQLAlchemy (\url{http://www.sqlalchemy.org}). This library was chosen
since it supports many SQL dialects. 
If SQLite is chosen, the database is stored as a single file which is
created automatically. This solution makes it possible to easily backup, copy and distribute
a database file. A server-based database, on the other hand, could be used by a collaboration 
who work together from different machines on the
same data: a central database server stores all data and the clients connect to
it to read or write data.

The database can store and manage all data that is retrieved via SunPy from the net
package though currently only the \textsc{VSO} package is supported. 
It also possible to add entries manually by file or directory. The package also provides
a \texttt{fetch} method which provides a unified data search which includes both local files
and files on the \textsc{VSO} are searched. This reduces the likelihood of downloading the same file 
multiple times. When a file is added to the database the file is scanned and a file hash is produced. 
The current date is associated with the entry along with metadata summaries such 
as instrument, date of observation, field of view, etc. 
The database also provides the ability to associate custom meta-data to 
each database entry such as keywords, comments, and favouring.

@@

\begin{listing}[H]
\begin{minted}[bgcolor=bg]{pycon}
>>> from sunpy.net import vso
>>> from sunpy.database import Database
>>> database = Database('sqlite:///')
>>> database.download(
...     vso.attrs.Time((2011, 9, 20, 1), (2011, 9, 20, 2)),
...     vso.attrs.Instrument('RHESSI'))
>>> len(database)
6
>>> from sunpy.database.tables import display_entries
>>> print display_entries(
...     database,
...     ['id', 'observation_time_start', 'wavemin', 'wavemax', 'download_time'])
id observation_time_start wavemin        wavemax           download_time      
-- ---------------------- -------        -------           -------------      
1  2011-09-20 01:09:20    0.413280643067 7.29318781883e-05 2014-01-18 19:21:35

### 2014-01-28 4-Util.tex
Some small typo fixed and latex stuff

\section{Additional functionality}
SunPy is meant to provide a consistent environment for solar data analysis. In 
order to achieve this goal SunPy provides a number of utility functions and packages which 
are used by the other SunPy modules and are made available to the user. This section 
briefly describes some of these functions.

@@

\subsection{World Coordinate System (WCS) Coordinates}
Coordinate transformations are frequently a necessary task within the solar 
data analysis workflow. Likely the most often used transformation is from 
observer coordinates (e.g. sky coordinates) to a coordinate system which is 
mapped onto the solar surface (e.g. latitude and longitude). This 
transformation is necessary to compare the true physical distance between 
different solar features. This type of transformation is not unique
to solar observations, but is not often considered by astronomical packages
such as the Astropy 
coordinates package. The \verb|wcs| package in SunPy implements the World Coordinate 
System (WCS) for solar coordinates as described by \cite{Thompson2000}. The 
transformations currently implemented are those most useful 
for most solar data analysis, namely converting from Helioprojective-Cartesian 
(HPC) to Heliographic coordinates (HG). HPC describes the positions on 
the Sun as angles measured from the center of the solar disk (usually in 
arcseconds) using cartesian coordinates (X, Y). This is the coordinate system 
most often defined in solar imaging data (see for example, images from SDO/AIA, SOHO/EIT, and TRACE). 
HG coordinates express positions on the Sun using longitude and latitude on 
the solar sphere. There are two standards for this coordinate system, 
Stonyhurst-Heliographic where the origin is at the intersection of the solar 
equator and the central meridian as seen from Earth and 
Carrington-Heliographic is a system which is fixed to the Sun and does not depend on Earth. The 
implementation of these transformations pass through a common coordinate system 
called Heliocentric-Cartesian (HCC) where positions are expressed in true 
(deprojected) physical distances instead of angles on the celestial sphere.
These transformations require some knowledge of the location of the observer 
which is usually provided by the image header. In the case where it is 
not, the observer is assumed to be at Earth.Listing \ref{code:wcs_code} shows some examples of coordinate transforms carried out in SunPy using the \verb|wcs| utilities.

@@

\subsubsection{Sun}
The purpose of this module is to provide solar-specific data for example ephemeris and
solar constants. The main namespace contains a number of functions which provide solar
ephemeris such as the sun to earth distance, solar cycle number, the mean anomaly, etc.
All of these functions take a time as their input which can be provided in any 
\texttt{parse\_time}
compatible format. 
A future version of SunPy will integrate with the AstroPy coordinates package and will 
likely make use of an ephemeris package.

\subsubsection{constants}
The purpose of the constants package is to provide a number of solar-related 
constants in order to provide consistency in the calculations of derived solar 
values both within the SunPy code base but also to the user. Every solar 
constant is provided as a Constant object as defined by Astropy. Each 
Constant object defines a Quantity, a number associated with a unit, along with 
the constant's provenance (i.e. reference) as well as its uncertainty. Using 
Astropy's Quantities, any solar constant can easily be converted between 
different units as can be seen in Figure~\ref{fig:constants_code} as well
as between the SI or cgs unit systems with a single command.  As these objects inherit from 
Numpy's ndarray, they work well with standard representations of numbers.
For convenience a number of shortcuts to frequently used constants are provided 
directly when importing the sub-package. A larger list of constants can be 
accessed through an interface modelled on that provided by the SciPy constants 
package and is available as a dictionary called \texttt{physical\_constants}. 
To view them all quickly a \texttt{print\_all()} function is available.

@@

\subsubsection{Instruments}
In addition to providing support for instrument-specific solar data via the main data 
classes \texttt{Map}, \texttt{LightCurve} and \texttt{Spectrum}, 
some instrument-specific functions may be found within the \texttt{instr} package. 
These functions are generally those that are unique to one particular solar instrument, 
rather than of general use, for example a function to construct a GOES flare event list 
or a function to query the LYRA timeline annotation file. Currently, some support is included
 in this way for the GOES, LYRA, RHESSI and IRIS instruments, while future developments 
 will include support for additional missions. It is anticipated that, ultimately, solar
  missions requiring a large suite of software tools will be supported via a separately 
  maintained package affiliated with SunPy.

>>>>

\section{Additional functionality}\label{sec:util}
SunPy is meant to provide a consistent environment for solar data analysis. In 
order to achieve this goal SunPy provides a number of utility functions and packages which 
are used by the other SunPy modules and are made available to the user. This section 
briefly describes some of these functions.

@@

\subsection{World Coordinate System (WCS) Coordinates}\label{ssec:util:wcs}
Coordinate transformations are frequently a necessary task within the solar 
data analysis workflow. Likely the most often used transformation is from 
observer coordinates (e.g. sky coordinates) to a coordinate system which is 
mapped onto the solar surface (e.g. latitude and longitude). This 
transformation is necessary to compare the true physical distance between 
different solar features. This type of transformation is not unique
to solar observations, but is not often considered by astronomical packages
such as the Astropy 
coordinates package. The \verb|wcs| package in SunPy implements the World Coordinate 
System (WCS) for solar coordinates as described by \cite{Thompson2000}. The 
transformations currently implemented are those most useful 
for most solar data analysis, namely converting from Helioprojective-Cartesian 
(HPC) to Heliographic coordinates (HG). HPC describes the positions on 
the Sun as angles measured from the center of the solar disk (usually in 
arcseconds) using cartesian coordinates (X, Y). This is the coordinate system 
most often defined in solar imaging data (see for example, images from SDO/AIA, SOHO/EIT, and TRACE). 
HG coordinates express positions on the Sun using longitude and latitude on 
the solar sphere. There are two standards for this coordinate system, 
Stonyhurst-Heliographic where the origin is at the intersection of the solar 
equator and the central meridian as seen from Earth and 
Carrington-Heliographic is a system which is fixed to the Sun and does not depend on Earth. The 
implementation of these transformations pass through a common coordinate system 
called Heliocentric-Cartesian (HCC) where positions are expressed in true 
(deprojected) physical distances instead of angles on the celestial sphere.
These transformations require some knowledge of the location of the observer 
which is usually provided by the image header. In the case where it is 
not, the observer is assumed to be at Earth.Listing \ref{code:wcs_code} shows some examples of coordinate transforms carried out in SunPy using the \verb|wcs| utilities.

@@

\subsubsection{Sun}\label{ssec:util:sun}
The purpose of this module is to provide solar-specific data for example ephemeris and
solar constants. The main namespace contains a number of functions which provide solar
ephemeris such as the sun to earth distance, solar cycle number, the mean anomaly, etc.
All of these functions take a time as their input which can be provided in any 
\texttt{parse\_time}
compatible format. 
A future version of SunPy will integrate with the AstroPy coordinates package and will 
likely make use of an ephemeris package.

\subsubsection{constants}\label{ssec:util:const}
The purpose of the constants package is to provide a number of solar-related 
constants in order to provide consistency in the calculations of derived solar 
values both within the SunPy code base but also to the user. Every solar 
constant is provided as a Constant object as defined by Astropy. Each 
Constant object defines a Quantity, a number associated with a unit, along with 
the constant's provenance (i.e. reference) as well as its uncertainty. Using 
Astropy's Quantities, any solar constant can easily be converted between 
different units as can be seen in Figure~\ref{fig:constants_code} as well
as between the SI or cgs unit systems with a single command.  As these objects inherit from 
Numpy's ndarray, they work well with standard representations of numbers.
For convenience a number of shortcuts to frequently used constants are provided 
directly when importing the sub-package. A larger list of constants can be 
accessed through an interface modelled on that provided by the SciPy constants 
package and is available as a dictionary called \texttt{physical\_constants}. 
To view them all quickly a \texttt{print\_all()} function is available.

@@

\subsubsection{Instruments}\label{ssec:util:inst}
In addition to providing support for instrument-specific solar data via the main data 
classes \texttt{Map}, \texttt{LightCurve} and \texttt{Spectrum}, 
some instrument-specific functions may be found within the \texttt{instr} package. 
These functions are generally those that are unique to one particular solar instrument, 
rather than of general use, for example a function to construct a GOES flare event list 
or a function to query the LYRA timeline annotation file. Currently, some support is included
 in this way for the GOES, LYRA, RHESSI and IRIS instruments, while future developments 
 will include support for additional missions. It is anticipated that, ultimately, solar
  missions requiring a large suite of software tools will be supported via a separately 
  maintained package affiliated with SunPy.

### 2014-01-28 Sunpy_paper_0.4.tex
added in my name and affiliation
Added affiliation
Some small typo fixed and latex stuff

\documentclass[12pt]{iopart}

%Uncomment next line if AMS fonts required
%\usepackage{iopams}

@@

\author{S J Mumford}
\address{Solar Physics \& Space Plasma Research Centre (SP$^{2}$RC), School of 
Mathematics and Statistics, The University of Sheffield, Hicks Building, 
Hounsfield Road, Sheffield, S3 7RH U.K.}

\author{J Bloggs}
\address{Wibble wiblle wiblle}

\ead{sunpy@googlegroups.com}

\begin{abstract}
This paper presents version 0.4 of SunPy a community developed Python package 
for solar physics.

\end{abstract}

@@

@@

%%% References
\bibliography{sunpy_paper_0.4}{}

>>>>

\documentclass[12pt]{iopart}
%Uncomment next line if AMS fonts required
%\usepackage{iopams}

\usepackage{natbib}
\newcommand*{\newblock}{Bibliography}

\usepackage{minted}
\definecolor{bg}{rgb}{0.95, 0.95, 0.95}
\usepackage{hyperref}

\usepackage{graphicx}
\graphicspath{{./images/}}

\usepackage{perpage} %the perpage package
\MakePerPage{footnote} %the perpage package command

@@

\input{authors}

\input{abstract}

@@

\input{solarsoft}

@@

%%% References
\bibliographystyle{stunat}
\bibliography{Sunpy_paper_0.4}{}
